---
title: "R Notebook"
output: html_notebook
---

Read in the mahoosive dataset and laod tidy

```{r}
library(tidyverse)
#dat=read.csv("C:/Users/ij926050/Dropbox/TREE/Data/Manziotta/NFI_Projections_CCMS_climate.csv")
dat


```
Visualise some of the data

ID = indvidual plot (within a trakt)
```{r}
ids=unique(dat$ID)
id_sample = sample(ids, 100)

dat%>%filter(ID==1)
dat1=dat%>%filter(PERIOD==1)
x=data.frame(table(dat1$ID,  dat1$ForestDomainData.TreatmentControlCategory))
hist(x$Freq)

dat%>%
 filter(ID %in% id_sample)%>%
  ggplot()+geom_point(aes(x=PERIOD, y = CarbonAndNitrogenData.Soil.Carbon.Stock
, col=ForestDomainData.TreatmentControlCategory

))+geom_line(aes(x=PERIOD, y = CarbonAndNitrogenData.Soil.Carbon.Stock
, col=ForestDomainData.TreatmentControlCategory

, group=ID))+facet_wrap(~region)+theme_bw()
```
Look at how many different schedules there are...

```{r}
man_dat = dat%>%filter(PERIOD==1)%>%dplyr::select(Description, ForestDomainData.TreatmentControlCategory)%>% 
     group_by(Description)%>%
  arrange(ForestDomainData.TreatmentControlCategory, .by_group = TRUE)%>%
       mutate(manage_options = paste0(ForestDomainData.TreatmentControlCategory, collapse = "__")) 
man_dat2=data.frame(table(man_dat$manage_options))

man_dat2%>%
  arrange(Freq, decreasing=T)

man_dat

```
```{r}
man_dat=man_dat%>%dplyr::select(Description, manage_options)%>%unique()

dat = merge(dat, man_dat, by="Description")


#this has a nice sample size
man_dat2=man_dat2%>%filter(Var1=="BAU FocusBioenergy__Lovgynnande trakthyggesbruk")
x=man_dat%>%filter(manage_options==man_dat2$Var1)


dat%>%filter(Description%in%x[1:6,]$Description)%>%ggplot()+geom_point(aes(x=PERIOD, y = CarbonAndNitrogenData.Soil.Carbon.Stock
, col=ForestDomainData.TreatmentControlCategory

))+geom_line(aes(x=PERIOD, y = CarbonAndNitrogenData.Soil.Carbon.Stock
, col=ForestDomainData.TreatmentControlCategory

, group=ForestDomainData.TreatmentControlCategory))+facet_wrap(~Description)+theme_bw()

```

```{r}
# take maximum age
dat2=dat%>%filter(manage_options=="BAU FocusBioenergy__Lovgynnande trakthyggesbruk")%>%slice_max(PERIOD)%>%filter(Climate=="Constant Climate")# take single year, single climate
dat2%>%filter(Description=="2008 1005 2180")
table(dat2$region);table(dat2$region2)

table(dat2$ForestDomainData.TreatmentControlCategory)


#recode as control and treatment and give nicer names
dat2=dat2%>%mutate(treatment = recode_factor(ForestDomainData.TreatmentControlCategory, "BAU FocusBioenergy" ="C", "Lovgynnande trakthyggesbruk" ="T"), soilc=CarbonAndNitrogenData.Soil.Carbon.Stock, age =ForestData.Mean.Age..excl.overstorey._Before, stems=ForestData.Stems_Before, temp =tsum_CC, rain=psum_CC)%>%
  dplyr::select(Description, ID,treatment, soilc,age,region, temp, rain, stems, BiomassData.Biomass.Total_Pine_Before:BiomassData.Biomass.Total_Birch_Before
, ForestData.Mean.Age..excl.overstorey._Before,  SiteData.Peat,SiteData.Soil.Moisture, TreatmentData.Biomass.Total.of.Cut.Trees.All.Species, SoilMoist, SoilMoist.cont)

```
EXtract age and stems at BEGINNING of managament period - before mgmt
```{r}
dat2%>%filter(Description=="2008 4552 4081")
datx=dat%>%filter(ID%in%dat2$ID)%>%slice_min(PERIOD)%>%filter(Climate=="Constant Climate")%>%group_by(ID) # take single year, single climate
datx=datx%>%ungroup%>%mutate(age.start =ForestData.Mean.Age..excl.overstorey._Before, stems.start=ForestData.Stems_Before, pine.start = BiomassData.Biomass.Total_Pine_Before
                             , spruce.start = BiomassData.Biomass.Total_Spruce_Before)%>%dplyr::select(Description, stems.start, age.start, pine.start, spruce.start)%>%unique()
datx$ID=NULL
datx%>%filter(Description=="2008 4552 4081")

dat2=merge(unique(dat2), unique(datx), "Description")
dat2
```





First step, need to assign a 'realised' management regime (treatment level) to each. 
Let's just go with random for now.

```{r}
dat3=dat2%>%
  group_by(Description)%>%
  slice_sample(n=1)

table(dat3$treatment)


dat3$treatment = as.numeric(dat3$treatment)-1
dat3


```

```{r}
custom_function = function(data, mapping, method = "loess", ...){
  p = ggplot(data = data, mapping = mapping) + 
    geom_point() + 
    geom_smooth(method=method, ...)
  
  p
}


dat1=data.frame(na.omit(dat3)%>%filter(treatment==1)%>%ungroup()%>%slice_sample(n=300)) # C only
dat0=data.frame(na.omit(dat3)%>%filter(treatment==0)%>%ungroup()%>%slice_sample(n=300)) # T only

dat1%>%dplyr::select(soilc,age.start, temp, rain,  stems.start,SoilMoist.cont, pine.start, spruce.start)%>% GGally::ggpairs(aes(alpha=0.5),progress=F,lower = list(continuous = custom_function))+theme_bw()
dat0%>%ungroup()%>%dplyr::select(soilc,age.start, temp, rain,  stems.start,SoilMoist.cont, pine.start, spruce.start)%>% GGally::ggpairs(aes(alpha=0.5),progress=F,lower = list(continuous = custom_function))+theme_bw()


#dat3%>%dplyr::select(soilc,age, temp, rain,  stems,SoilMoist.cont)%>% GGally::ggpairs(aes(alpha=0.5, col=treatment),progress=F,lower = list(continuous = custom_function))+theme_bw()



```
the distributions of covariates are similar for both treatments, because  I allocated them randomly!


```{r}
library(ranger)

n_features = 6

n.points=round(1*nrow(dat0))
#dat0.train=data.frame(dat0%>%slice_sample(n=n.points))
# train a default random forest model
m.0_rf1 <- ranger(
  formula  = soilc ~ age.start+  temp +  rain +  stems.start + SoilMoist, 
  data = dat2,
#  mtry = floor(n_features / 3),
  respect.unordered.factors = "order",
  seed = 123
)

# get OOB RMSE
(default_rmse <- sqrt(m.0_rf1$prediction.error))
## [1] 24859.27

# create hyperparameter grid
hyper_grid <- expand.grid(
  mtry = floor(n_features * c(.05, .15, .25, .333, .4)),
  min.node.size = c(1, 3, 5, 10), 
  replace = c(TRUE, FALSE),                               
  sample.fraction = c(.5, .63, .8),                       
  rmse = NA                                               
)

#hyper_grid <- hyper_grid[sample(nrow(hyper_grid),10),]

# execute full cartesian grid search
for(i in seq_len(nrow(hyper_grid))) {
  # fit model for ith hyperparameter combination
  fit <- ranger(
  formula  = soilc ~ age+ temp +  rain +  stems + SoilMoist.cont, 
    data            = data.frame(dat0), 
    num.trees       = n_features * 10,
    mtry            = hyper_grid$mtry[i],
    min.node.size   = hyper_grid$min.node.size[i],
    replace         = hyper_grid$replace[i],
    sample.fraction = hyper_grid$sample.fraction[i],
    verbose         = FALSE,
    seed            = 123,
    respect.unordered.factors = 'order',
  )
  # export OOB error 
  hyper_grid$rmse[i] <- sqrt(fit$prediction.error)
  #hyper_grid$mtry[i] <- fit$mtry

}


# assess top 10 models
hyper_grid %>%
  arrange(rmse) %>%
  mutate(perc_gain = (default_rmse - rmse) / default_rmse * 100) %>%
  head(10)

hyper_grid %>%ggplot()+geom_jitter(aes(x=min.node.size, y=rmse, col=mtry, shape=as.factor(sample.fraction)))+theme_bw()+facet_wrap(~replace)

```






```{r}
dat.all=rbind(dat1,dat0)
dat.all%>%ggplot()+geom_density(aes(soilc, fill=as.factor(treatment)),alpha=0.5)
dat.all%>%ggplot()+geom_density(aes(soilc, fill=as.factor(treatment)),alpha=0.5)+facet_wrap(~cut(SoilMoist.cont,2))
#dat.all$treatment=as.numeric(dat.all$treatment)
#dat.all$treatment=as.numeric(dat.all$treatment)-1

mall.rf_impurity <- ranger(
  formula  = soilc ~ treatment + temp +  rain +  stems.start + age.start +spruce.start+pine.start+ SoilMoist.cont, 
  data = data.frame(dat.all), 
  num.trees = 2000,
  mtry = 3,
  min.node.size = 4,
  sample.fraction = 1,
  replace = FALSE,
  importance = "impurity",
 # respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)

p1.0 <- vip::vip(mall.rf_impurity, num_features = 8, geom="point")+theme_bw()



m1.rf_impurity <- ranger(
  formula  = soilc ~  temp +  rain +  stems.start + age.start + SoilMoist.cont+spruce.start+pine.start, 
  data = dat1, 
  num.trees = 2000,
  mtry = 1,
  min.node.size = 2,
  sample.fraction = .80,
  replace = FALSE,
  importance = "impurity",
 # respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)


m1.rf_permutation <- ranger(
  formula  = soilc ~  temp +  rain +  stems.start + age.start + SoilMoist.cont+spruce.start+pine.start, 
  data = dat1, 
  num.trees = 2000,
  mtry = 1,
  min.node.size = 2,
  sample.fraction = 1,
  replace = FALSE,
  importance = "permutation",
 # respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)

p1.1 <- vip::vip(m1.rf_impurity, num_features = 6, geom="point")+theme_bw()
p2 <- vip::vip(m1.rf_permutation, num_features = 6, bar = FALSE)
gridExtra::grid.arrange(p1.1, p2, nrow = 1)


m0.rf_impurity <- ranger(
  formula  = soilc ~  temp +  rain +  stems.start + age.start + SoilMoist.cont+spruce.start+pine.start, 
  data = dat0, 
  num.trees = 2000,
  mtry = 1,
  min.node.size = 2,
  sample.fraction = 1,
  replace = FALSE,
  importance = "impurity",
  #respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)


m0.rf_permutation <- ranger(
  formula  = soilc ~  temp +  rain +  stems.start + age.start + SoilMoist.cont+spruce.start+pine.start, 
  data = dat0, 
  num.trees = 2000,
  mtry = 1,
  min.node.size = 2,
  sample.fraction = 1,
  replace = FALSE,
  importance = "permutation",
 # respect.unordered.factors = "order",
  verbose = FALSE,
  seed  = 123
)

p1.2 <- vip::vip(m0.rf_impurity, num_features = 8, geom="point")+theme_bw()
p2 <- vip::vip(m0.rf_permutation, num_features = 8, bar = FALSE)
gridExtra::grid.arrange(p1.2, p2, nrow = 1)

gridExtra::grid.arrange(p1.0, p1.1,p1.2, nrow = 1)
```

Generate test data 
```{r}
test.data=dat3%>%filter(!Description %in% rbind(dat1$Description,dat0$Description))# remove the sampled training points
```

```{r}

predC.slearn.0 = predict(mall.rf_impurity, data=test.data%>%mutate(treatment=0))
test.data$predC.slearn.0 <- predC.slearn.0$predictions

predC.slearn.1 = predict(mall.rf_impurity, data=test.data%>%mutate(treatment="T"))
test.data$predC.slearn.1 <- predC.slearn.1$predictions


predC.tlearn.0 = predict(m0.rf_impurity, data=test.data)
test.data$predC.tlearn.0 <- predC.tlearn.0$predictions

predC.tlearn.1 = predict(m1.rf_impurity, data=test.data)
test.data$predC.tlearn.1 <- predC.tlearn.1$predictions

test.data=test.data%>%mutate(ITE_tlearn = predC.tlearn.1-predC.tlearn.0)
test.data=test.data%>%mutate(ITE_slearn = predC.slearn.1-predC.slearn.0)

#now lets compute the true ITEs
#test data has just one treatment per individual

true=dat2%>%dplyr::select(Description, treatment, soilc)%>%unique()%>%pivot_wider(id_cols = Description,values_from = "soilc", names_from=treatment)%>%na.omit()%>%mutate(true_ITE= T-C)
test.data=merge(test.data, true%>%dplyr::select(Description, true_ITE), by="Description")



test.data%>%group_by(Description)

test.data%>%ggplot()+geom_point(aes(x=treatment, y=soilc))+geom_line(aes(group=Description,x=treatment, y=soilc))

test.data%>%ggplot()+geom_point(aes(x=true_ITE, y=ITE_tlearn))+geom_line(aes(x=true_ITE, y=true_ITE))

test.data%>%ggplot()+geom_point(aes(x=true_ITE, y=ITE_slearn))+geom_line(aes(x=true_ITE, y=true_ITE))

hist(test.data$true_ITE)



px1=test.data%>%filter(treatment=="T")%>%ggplot()+geom_point(aes(x=soilc, y=predC.tlearn.1))+facet_grid(~treatment)+geom_line(aes(x=soilc, y=soilc))+ylim(30,125)+xlim(30,125)
px2=test.data%>%filter(treatment=="C")%>%ggplot()+geom_point(aes(x=soilc, y=predC.tlearn.0))+facet_grid(~treatment)+geom_line(aes(x=soilc, y=soilc))+ylim(30,125)+xlim(30,125)
px3=test.data%>%ggplot()+geom_point(aes(x=soilc, y=predC.slearn))+facet_grid(~treatment)+geom_line(aes(x=soilc, y=soilc))+ylim(30,125)+xlim(30,125)

#gridExtra::grid.arrange(px1,px2,px3, ncol=3, widths=c(1,1,2))


test.data%>%filter(treatment==1)%>%ggplot()+geom_point(aes(x=predC.slearn.1, y=predC.tlearn.1))+facet_grid(~treatment)+geom_line(aes(x=soilc, y=soilc))

test.data%>%filter(treatment==0)%>%ggplot()+geom_point(aes(x=predC.slearn.0, y=predC.tlearn.0))+facet_grid(~treatment)+geom_line(aes(x=soilc, y=soilc))

```



```{r}
true=dat2%>%filter(Description %in% test.data$Description)%>%dplyr::select(Description, treatment, soilc)%>%unique()%>%pivot_wider(id_cols = Description,values_from = "soilc", names_from=treatment)%>%na.omit()%>%mutate(true_ITE= T-C)

#slearn=test.data%>%dplyr::select(Description, treatment, predC.slearn)%>%pivot_wider(id_cols = Description,values_from = "predC.slearn", names_from=treatment)%>%na.omit()%>%mutate(slearn_ITE=T-C)

#tlearn=test.data%>%dplyr::select(Description, treatment, predC.tlearn.1, predC.tlearn.0)%>%pivot_wider(id_cols = Description,values_from = "predC.tlearn.1", names_from=treatment)%>%na.omit()%>%mutate(slearn_ITE=T-C)






test.data=test.data%>%mutate(tlearn_ITE=predC.tlearn.1-predC.tlearn.0)
test.data=test.data%>%mutate(slearn_ITE=predC.slearn.1-predC.slearn.0)

test.data  <- merge(test.data, true, "Description")
#test.data  <- merge(test.data, slearn%>%dplyr::select(Description, slearn_ITE), "Description")

test.data%>%dplyr::select(Description, treatment, soilc, predC.slearn.0,predC.slearn.1, predC.tlearn.0, predC.tlearn.1, ITE_tlearn, slearn_ITE, true_ITE.y)

p1=test.data%>%filter(soilc>0)%>%ggplot(aes(x=true_ITE.y, y=tlearn_ITE, col=soilc))+geom_point()+geom_line(aes(x=true_ITE.y, y=true_ITE.y))+theme_bw()+ylab("ITE prediction, \nT-learner approach")+xlab("True ITE")#+ylim(-120,100)
p2=test.data%>%filter(soilc>0)%>%ggplot(aes(x=true_ITE.y, y=slearn_ITE,col=soilc))+geom_point()+geom_line(aes(x=true_ITE.y, y=true_ITE.y))+theme_bw()+ylab("ITE prediction, \nS-learner approach")+xlab("True ITE")#+ylim(-120,100)
p3=test.data%>%filter(soilc>0)%>%ggplot(aes(x=tlearn_ITE, y=slearn_ITE,col=soilc))+geom_point()+geom_line(aes(x=true_ITE.y, y=true_ITE.y))+theme_bw()+ylab("ITE prediction, \nS-learner approach")+xlab("ITE prediction, \nT-learner approach")#+ylim(-120,100)


gridExtra::grid.arrange(p1, p2, p3, nrow=1)
```



```{r}
library(randomForest)
test.data=dat3%>%filter(!Description %in% rbind(dat1$Description,dat0$Description))# remove the sampled training points



varsforcethin <- test.data
varsforcethin$treatment=1
varsforceunthin <- test.data
varsforceunthin$treatment=0

library(randomForest)
thinRF <- randomForest(x = dat1%>%dplyr::select(temp,  rain,  stems.start, age.start, SoilMoist.cont), y = as.numeric(dat1$soilc),  
                  method = "rf", metric = "ROC",  ntree=1000,
               importance = TRUE, stepFactor=1.5,keep.inbag=TRUE)

unthinRF <- randomForest(x = dat0%>%dplyr::select(temp,  rain,  stems.start, age.start, SoilMoist.cont), y = as.numeric(dat0$soilc),  
                  method = "rf", metric = "ROC",  ntree=1000,
               importance = TRUE, stepFactor=1.5)

allRF <- randomForest(x = dat.all%>%dplyr::select(treatment,temp,  rain,  stems.start, age.start, SoilMoist.cont), y = as.numeric(dat.all$soilc),  
                  method = "rf", metric = "ROC",  ntree=1000,
               importance = TRUE, stepFactor=1.5)

importance(allRF)
#plot(margin(thinRF)) 
par(mfrow=c(2,3))
#for(i in 1:6) partialPlot(thinRF,varsthinned,x.var=names(varsthinned%>%dplyr::select(!madake))[i])
#for(i in 1:6) partialPlot(unthinRF,varsunthinned,x.var=names(varsthinned%>%dplyr::select(!madake))[i])

#library(plotmo)
#vars$meantemp
#plotmo(thinRF,"meantemp", type="prob")

test.data$predDIFF.RF <- predict(unthinRF, newdata=varsforceunthin, type="response")-predict(thinRF, newdata=varsforcethin, type="response")
hist(test.data$predDIFF.RF)

#vars2$predDIFF.glm <- predict(bestsecdecid, newdata=vars2%>%mutate(canopy_cover3="unthinned"), type="response")-predict(bestsecdecid, newdata=vars2%>%mutate(canopy_cover3="thinned"), type="response")


#vars2$predDIFF.glm <- predict(bestsecdecid, newdata=vars2%>%mutate(canopy_cover3="unthinned"), type="response")-predict(bestsecdecid, newdata=vars2%>%mutate(canopy_cover3="thinned"), type="response")

test.data$predDIFF.allRF <- predict(allRF, newdata=varsforcethin, type="response")-predict(allRF, newdata=varsforceunthin, type="response")

test.data%>%ggplot()+geom_point(aes(x=predDIFF.RF, y=predDIFF.allRF))
merge(test.data, true%>%dplyr::select(Description, true_ITE))%>%
  ggplot()+geom_point(aes(x=true_ITE, y=predDIFF.allRF))

merge(test.data, true%>%dplyr::select(Description, true_ITE))%>%
  ggplot()+geom_point(aes(x=true_ITE, y=predDIFF.RF))
```


