---
title: "Exploring canopy cover"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("ggcorrplot")
```

Since splitting plots into thinning/no-thinning groups doesn't seem to give us very good sample sizes,
we are investigating using a continuous variable for our treatment effects. 

I'm going to take a quick look at canopy cover as one option.

```{r message=FALSE}
read_delim(here::here("data", "raw", "nfi_plots_sp.txt")) -> plot_data

read_delim(here::here("data", "raw", "nfi_f_mactions.txt")) -> management_data

read_delim(here::here("data", "raw", "nfi_climat.txt")) -> climate_data
```

```{r}
# filter data for only land class 1 and 1st year of data
plot_data %>% 
  filter(InternationalCodeLandUse == 1) %>% 
  group_by(ID) %>% 
  slice_min(Year) %>%
  ungroup() %>% 
  rename(UniqueID = uniqueID) -> filtered_plots

```

```{r}
filtered_plots %>% 
  ggplot(aes(x = CanopyCover)) +
  geom_density()

```

## Pine dominated forests

We still want to compare forests that are fairly similar so let's look at pine-dominated forests only.

```{r}
filtered_plots %>% 
  ggplot(aes(x = PineVol)) +
  geom_histogram() 

```

Calculate volume of pine as a percentage.

```{r}
filtered_plots %>%
  rowwise() %>% 
  mutate(percent_pine = (PineVol / Volume)*100) -> pine_plots
  
pine_plots %>% 
  ggplot(aes(x = percent_pine)) +
  geom_histogram() 
```

```{r}
pine_plots %>% 
  filter(percent_pine >= 50) %>% 
  n_distinct("ID")

```

Let's say that plots with greater than 50% volume Pine are Pine dominated forests. 
That would give us a sample size of 2,845 plots to work with.

```{r}
pine_plots %>% 
  filter(percent_pine > 50) %>% 
  ggplot(aes(x = CanopyCover)) +
  geom_density()

```

## Correlations

How does canopy cover correlate with other variables?

```{r}
climate_data %>% 
  group_by(ID, year) %>% 
  summarise(avg_temp = mean(Temperature),
            avg_precip = mean(Precipitation)) %>% 
  ungroup() %>% 
  rename(Year = year) %>% 
  inner_join(pine_plots, by = c("ID", "Year")) -> pine_plots_cl
```

```{r warning=FALSE}
pine_plots_cl %>% 
  drop_na(CanopyCover) %>% 
  select(CanopyCover, BasalArea, Volume, NrStems, SlopeDirection, StandAge, 
         StandHeight, avg_temp, avg_precip) %>% 
  mutate_if(is.numeric, round) %>% 
  cor(method = "spearman", use = "pairwise.complete.obs") -> cor_mat

ggcorrplot(cor_mat, hc.order = TRUE, type = "lower",
     outline.col = "white", lab = TRUE)
```

Canopy cover correlates positively with `NrStems`, `StandHeight`, `BasalArea` and `Volume`,
which makes sense as they're all measuring similar things.

The correlations with `StandAge` and `avg_temp` might be worth looking into a fit further.

```{r warning=FALSE}
pine_plots_cl %>% 
  ggplot(aes(x = CanopyCover, y = avg_temp)) +
  geom_point() +
  geom_smooth(method = "lm")

```

Looks like we almost have 2 clusters here, maybe surveys taken in the same year/adjacent years?

```{r warning=FALSE}
pine_plots_cl %>% 
  ggplot(aes(x = CanopyCover, y = StandAge)) +
  geom_point() +
  geom_smooth(method = "lm")

```

This one seems counter-intuitive to me?

# thinning

```{r}
# filter management data for only 1st survey year
# and join to plot data by the plot x yr variable
management_data %>% 
  group_by(ID) %>% 
  slice_min(Year) %>%
  ungroup() -> man_data

```

```{r}
# get timing of management action
man_data %>% 
  mutate(TimeAction = str_remove(TimeAction, "\xc5r/s\xe4song ")) %>%
  mutate(MedYrAgo = case_when(TimeAction == "0" ~ 0,
                              TimeAction == "1" ~ 1,
                              TimeAction == "2" ~ 2,
                              TimeAction == "3-5" ~ 4,
                              TimeAction == "6-10" ~ 8,
                              TimeAction == "11-25" ~ 18,
                              TimeAction == "26+" ~ 26)) -> man_data_time
```

```{r}
# order and concatenate 
man_data_time %>% 
  group_by(ID) %>% 
  arrange(MedYrAgo, .by_group = TRUE) %>%
  summarise(CodeAction = str_c(CodeAction, collapse = ", "), 
            .groups = "drop") ->  man_hists

```

```{r}
paste0(c(20:23, 10:13, 30:33, 40, 43:46)) %>% 
  str_c(collapse = "|") -> all_thin

```

```{r}
man_hists %>% 
  filter(
      str_detect(CodeAction, all_thin)
  ) -> thin_plots
  
pine_plots_cl %>% 
  mutate(thinned = case_when(
    ID %in% thin_plots$ID ~ TRUE,
    .default = FALSE)) %>% 
  drop_na(thinned) -> thin_pine
```

```{r warning=FALSE}
thin_pine %>% 
  ggplot(aes(x = CanopyCover, y = StandAge, colour = thinned)) +
  geom_point() +
  geom_smooth(method = "lm")

```

```{r warning=FALSE}
thin_pine %>% 
  ggplot(aes(x = CanopyCover, y = avg_temp, colour = thinned)) +
  geom_point() +
  geom_smooth(method = "lm")

```

Wow that's cool, thinned plots are warmer!
