---
title: "Causal quartets"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

_Try and make some nice intuitive figures that clearly show how ATEs differ from ITEs. Or how the same ATE can give a very different distribution and spread of ITEs._

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("patchwork")
```

## `rlearner` package
[xnie/rlearner](https://rdrr.io/github/xnie/rlearner/src/R/utils.R): R-learner for Heterogeneous Treatment Effect Estimation

The `easy_toy_data_simulation` function generates a toy dataset of size `n` that can be used to experiment with the learners and meta-learners.

output: a list containing the covariate matrix (`x.covariate_1` and `x.covariate_2`), treatment vector `w`, outcome vector `y`, true propensity vector `p`, true marginal outcome vector `m`, true control potential outcome vector `mu0`, true treated potential outcome vector `mu1`, and true treatment effect vector `tau`, in that order.

```{r}
easy_toy_data_simulation = function(n) {
	x = stats::model.matrix(~.-1, data.frame("covariate_1" = rnorm(n), "covariate_2"= rnorm(n)))
	p = rep(0.5, n)
	w = as.numeric(rbinom(n,1,p)==1)
	tau = x %*% c(1, 1)
	m = x %*% c(0.5, -0.5)
	mu1 = m + tau/2
	mu0 = m - tau/2
	y = (m + tau/2*(2*w-1))[,1]
	list(x=x, w=w, y=y, p=p, m=m, mu0=mu0, mu1=mu1, tau=tau)
}
```

```{r}
easy_toy_data_simulation(1000) %>% 
  as.data.frame() %>% 
  ggplot(aes(x = x.covariate_1, y = y, colour = as.factor(w))) + 
  geom_point()
```

## Blog post

Code from a blog post about the `personalized` package:

https://cran.r-project.org/web/packages/personalized/vignettes/efficiency_augmentation_personalized.html

```{r}
set.seed(1)
n.obs  <- 1000
n.vars <- 10
x <- matrix(rnorm(n.obs * n.vars, sd = 3), n.obs, n.vars)

# simulate non-randomized treatment
xbetat   <- 0.5 + 0.25 * x[,9] - 0.25 * x[,1]
trt.prob <- exp(xbetat) / (1 + exp(xbetat))
trt      <- rbinom(n.obs, 1, prob = trt.prob)

# simulate delta
delta <- (0.5 + x[,2] - 0.5 * x[,3] - 1 * x[,1] + 1 * x[,1] * x[,4] )

# simulate main effects g(X)
xbeta <- 2 * x[,1] + 3 * x[,4] - 0.25 * x[,2]^2 + 2 * x[,3] + 0.25 * x[,5] ^ 2
xbeta <- xbeta + delta * (2 * trt - 1)

# simulate continuous outcomes
y <- drop(xbeta) + rnorm(n.obs, sd = 3)

# make data frame
tibble(xbeta, y, trt = as.factor(trt), delta) -> sim_dat
```

```{r}
sim_dat %>% 
  ggplot(aes(x = xbeta, y = y, colour = trt)) + 
  geom_point()
```

```{r}
sim_dat %>% 
  ggplot(aes(x = xbeta, y = delta, colour = trt)) + 
  geom_point()
```

## `quartets` package

Contains the data from [Causal quartets: Different ways to attain the same average treatment effect](http://www.stat.columbia.edu/~gelman/research/unpublished/causal_quartets.pdf)

A dataframe with 88 rows and 5 variables:

-   dataset: The data generating mechanism 
-   exposure: exposure
-   covariate: a pre-exposure factor
-   outcome: outcome
-   .causal_effect: latent true causal effect

Is the `.causal_effect` the individual treatment effect? 
i.e. the difference between the effects with and without the treatment?
The `covariate` is `x`, the thing you are using to try and predict the outcome,
So to see the ATE you'd take the average `.causal_effect` from each group (`exposure` yes or no)?

```{r}
quartets::heterogeneous_causal_quartet %>%  glimpse()
```

```{r}
quartets::heterogeneous_causal_quartet %>%
  ggplot(aes(x = covariate, y = outcome, colour = as.factor(exposure))) + 
  geom_point() +
  geom_smooth(method = lm) -> p1

quartets::heterogeneous_causal_quartet %>%
  ggplot(aes(x = outcome, colour = as.factor(exposure))) + 
  geom_density() -> p2

p1 + p2 + plot_layout(guides = "collect")
```

So say you ran an experiment and got this data. 
You'd think that exposure to the treatment gave a better outcome.

But these are average treatment effects...

These 4 plots below have the same average effects but different individual effects.

```{r}
quartets::heterogeneous_causal_quartet %>%
  ggplot(aes(x = covariate, y = outcome, colour = as.factor(exposure))) + 
  geom_point() +
  facet_wrap(~dataset)
```


```{r}
quartets::heterogeneous_causal_quartet %>%
  ggplot(aes(x = outcome, colour = as.factor(dataset))) + 
  geom_density() 
```
