---
title: "Find treatment groups"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

We need to identify treatment and control groups from the Swedish NFI data. 
For example, thinned vs non-thinned stands. 
Ideally identify stands with similar management histories/thinning histories,
but vary in whether they have been recently thinned at a particular stage.

The [Swedish NFI handbook](https://www.slu.se/globalassets/ew/org/centrb/rt/dokument/faltinst/nfi_fieldwork_instructions_eng.pdf) 
has details about how the data was collected.

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
```

```{r message=FALSE}
read_delim(here::here("data", "raw", "nfi_plots_sp.txt")) -> plot_data

read_delim(here::here("data", "raw", "nfi_f_mactions.txt")) -> management_data

read_delim(here::here("data", "raw", "nfi_climat.txt")) -> climate_data
```

## Check what the different IDs mean

Tord has told us that `UniqueID` is the unique id for each plot x year sampling event.
`PlotNr` looks like it could be a unique plot id. 
We can check this by seeing if there are approx 3 years per plot 
(each plot should have been sampled approximatley 3 times).

```{r}
plot_data %>% 
  summarise(n_years = n_distinct(Year), .by = "PlotNr")

```

Ok, there are 29 or 28 entries per `PlotNr` so that can't be a unique plot ID, 
let's try `ID`

```{r}
plot_data %>% 
  summarise(n_years = n_distinct(Year), .by = "ID")

```

Yep, that looks better. 3 years per `ID`. Maybe we could use years to split the data up into test and train.

Let's check that we only have 1 year for each of the `uniqueID` values.

```{r}
plot_data %>% 
  summarise(n_years = n_distinct(Year), .by = "uniqueID")

```

Sweet, all good!

## Compare plots with similar management histories

We want our control and treatment groups to be as similar as possible except in what the treatment actually is. I might just go really broad for now and use 'land classes' to pick plots that are relatively similar.

International land use classes:

1. Forest
2. Other wooded land
3. Bare impediment

For snapshot sampling we can use the first year from the permanent plots and all of the temporary plots (which are only sampled once). 
Lets filter for that too.

```{r}
# filter data for only land class 1 and 1st year of data
plot_data %>% 
  filter(InternationalCodeLandUse == 1) %>% 
  group_by(ID) %>% 
  slice_min(Year) %>%
  ungroup() %>% 
  rename(UniqueID = uniqueID) -> filtered_plots

```

We need to find plots with similar management histories. 
Let's bring in the management data. 

```{r}
# filter management data for only 1st survey year
# and join to plot data by the plot x yr variable
management_data %>% 
  group_by(ID) %>% 
  slice_min(Year) %>%
  ungroup() %>%
  mutate(UniqueID = gsub("}", "", UniqueID)) %>% 
  mutate(UniqueID = gsub("\\{", "", UniqueID)) %>% 
  left_join(filtered_plots) -> plt_man_data

```

Ok, so the `TimeAction` variable gives us information about when the management action happened.
It's in an awkward format so let's fix that.

```{r}
plt_man_data %>% 
  mutate(TimeAction = str_remove(TimeAction, "\xc5r/s\xe4song ")) %>%
  mutate(MedYrAgo = case_when(TimeAction == "0" ~ 0,
                              TimeAction == "1" ~ 1,
                              TimeAction == "2" ~ 2,
                              TimeAction == "3-5" ~ 4,
                              TimeAction == "6-10" ~ 8,
                              TimeAction == "11-25" ~ 18,
                              TimeAction == "26+" ~ 26)) -> man_data_time
```

There might be common patterns of management, 
e.g. they might often do some sapling clearing in the first year followed by thinning in the 10th year. 

If we order the actions by when they happened for each plot (using the `MedYrAgo` variable we created above), 
and concatenate into one column we can look for common patterns.

```{r}
man_data_time %>% 
  group_by(ID) %>% 
  arrange(MedYrAgo, .by_group = TRUE) %>%
  summarise(CodeAction = str_c(CodeAction, collapse = ", "), 
            .groups = "drop") ->  man_hists

man_hists %>%   
  summarise(n = n(), .by = CodeAction) %>% 
  arrange(-n) 
```

We're looking for plots with a similar history but differences in their more recent management.
22 seems quite common and I can see some patterns of 22 followed by: 22 again (78 plots), 43 (24 plots), 21 (21 plots) and 31 (20 plots).
31 and 43 could be good candidates too. 

```{r}
filtered_plots %>% 
  inner_join(man_hists) %>% 
  filter(grepl("^22", CodeAction)) %>% 
  filter(CodeAction == "22" |
           CodeAction == "22, 22" |
           CodeAction == "22, 43" |
           CodeAction == "22, 21" |
           CodeAction == "22, 31") %>% 
  ggplot(aes(x = `Cover_Vaccinium myrtillus`, colour = CodeAction, fill = CodeAction)) +
  geom_density(alpha = 0.3)

```

Let's make a plotting function and try this out for a few different variables.

```{r}
plot_man_hist <- function(data, plot_var) {
  data %>% 
  filter(grepl("^22", CodeAction)) %>% 
  filter(CodeAction == "22" |
           CodeAction == "22, 22" |
           CodeAction == "22, 43" |
           CodeAction == "22, 21" |
           CodeAction == "22, 31") -> filtered_dat 
  ggplot(filtered_dat, aes(x = .data[[plot_var]], colour = CodeAction, fill = CodeAction)) +
  geom_density(alpha = 0.3)
}

filtered_plots %>% 
  inner_join(man_hists) -> plots_hist_data

var_list <- list("StandAge", "CanopyCover", "PineVol", "AspenVol",
                 "NorwaySpruceVol", "BirchVol")

plot_list <- lapply(var_list, plot_man_hist, data = plots_hist_data)

patchwork::wrap_plots(plot_list) + 
  patchwork::plot_layout(ncol = 2)

```

## Compare 22 ("other thinning") to non-thinned plots

Let's try to group plots into thinned and non-thinned.

```{r}
paste0(c(20:23, 10:13, 30:33, 40, 43:46)) %>% 
  str_c(collapse = "|") -> all_thin

# get list of plots that have the 22 management action and no others
man_hists %>% 
  filter(CodeAction == "22") %>%  
  pull(ID) -> plots_22

```

If thinning is recorded for a plot at any time point I'm classing it as thinned.

```{r}
filtered_plots %>% 
  filter(ID %in% plots_22) -> thin_plots_22

man_hists %>% 
  filter(
      str_detect(CodeAction, all_thin, negate = TRUE)
  ) -> no_thin_plots
  
filtered_plots %>% 
  mutate(thinned = case_when(
    ID %in% plots_22 ~ TRUE,
    ID %in% no_thin_plots$ID ~ FALSE,
    .default = NA)) %>% 
  drop_na(thinned) -> thin_no_thin
```

I'm assuming that NAs mean we don't have management data for these plots rather than not having any management happen to them. 
I think that's right... 
Have dropped them here but it takes nrows down from 6,730 to 587. 
That's a lot of missing data!?

There are `r n_distinct(thin_plots_22$ID)` thinned plots (action 22 only) and `r n_distinct(no_thin_plots$ID)` plots which have not been thinned at all.

Compare a few variables in plots with and without thinning.

```{r}
plot_man_thin <- function(data, plot_var) {
  thin_no_thin %>% 
  ggplot(aes(x = .data[[plot_var]], colour = thinned, fill = thinned)) +
  geom_density(alpha = 0.3)
}

var_list_2 <- list("StandAge", "StandHeight", "CanopyCover", "PineVol", 
                 "NorwaySpruceVol", "BirchVol", "Cover_Vaccinium vitis-idaea", 
                 "Cover_Vaccinium myrtillus")

plot_list_2 <- lapply(var_list_2, plot_man_thin, data = thin_no_thin)

patchwork::wrap_plots(plot_list_2) + 
  patchwork::plot_layout(guides = 'collect')

```

## Climate data

Climate data is by month so I'll take an average temp and precipitation per plot per year.

```{r}
climate_data %>% 
  group_by(ID, year) %>% 
  summarise(avg_temp = mean(Temperature),
            avg_precip = mean(Precipitation)) %>% 
  ungroup() %>% 
  rename(Year = year) %>% 
  inner_join(thin_no_thin, by = c("ID", "Year")) -> thin_no_thin_cl
```

First looking at temperature

```{r}
ggplot(thin_no_thin_cl, aes(x = avg_temp, y = `Cover_Vaccinium myrtillus`, 
                            colour = thinned)) +
  geom_point(alpha = 0.7) +
  facet_grid(~thinned)

```

Now precipitation

```{r}
ggplot(thin_no_thin_cl, aes(x = avg_precip, y = `Cover_Vaccinium myrtillus`, 
                            colour = thinned)) +
  geom_point(alpha = 0.7) +
  facet_grid(~thinned)

```

It feels like the no-thin sample size is too small to see any patterns..
