---
title: "Explore new data"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

Tord has been able to send us the other 4/5th of the data :tada:

Let's take a look at it and compare it to the data we already had.

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("janitor")
```

Load new data

```{r}
file_names <- 
  as.list(dir(path = here::here("data", "raw"),
                         pattern = "ForManSims*", full.names = TRUE))

# don't load the first big file we recieved
file_names[[1]] <- NULL

data_list <- 
  lapply(X = file_names, FUN = read.csv, sep = ";", skipNul = TRUE)


names(data_list) <- 
  lapply(file_names, basename)

new_data <- 
  bind_rows(data_list, .id='df') %>%
  clean_names() 

rm(data_list)
glimpse(new_data)  
```

Load original data

```{r}
og_data <-
  readRDS(here::here("data", "derived", "subset_ForManSims_2016_RCP0.rds")) %>%
  mutate(df = "ForManSims_2016_RCP0.csv")
```

## Do the new data contain new plots?

The new data contains `r n_distinct(new_data$description)` unique plots (`description` IDs), 
and the original data has `r n_distinct(og_data$description)`.

If we join the data does the number of unique plots = 30057 (25326 + 4731)?

```{r}
og_data_0 <- 
  og_data %>% 
  filter(period == 0)

new_data_0 <- 
  new_data %>% 
  filter(period == 0)

bind_rows(og_data_0, new_data_0) %>% 
  distinct(description) %>% 
  glimpse()
  
```

We get 29,939 distinct plots, so there's an overlap of 118 plots. 
Where are they duplicated?

```{r}
inner_join(og_data_0, new_data_0, by = "description") %>% 
  select(df.x, df.y)
```

The overlaps are between `ForManSims_2017_RCP0.csv` and `ForManSims_2016_RCP0.csv` (the original data).

Are all the variables duplicated or just the plot ID?

```{r}
bind_rows(og_data_0, new_data_0) %>% 
  select(-df) %>% 
  get_dupes() %>% 
  nrow()
```

Looks like the entire rows are duplicated (236/2=118). 
That's ok, these are easy to remove. 

```{r}
bind_rows(og_data_0, new_data_0) %>% 
  distinct(across(-df), .keep_all = TRUE) -> all_data_0
```

## What management options have been simulated for plots in the new data?

```{r}
summary(as.factor(new_data$control_category_name))
```

Just the ones we want :thumbsup:

## What is our sample size?

Keep getting the `Error: vector memory exhausted (limit reached?)` message,
might have to test this on the cluster.
