---
title: "Get climatic variables for NFI plots"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("janitor")
library("tidymodels")
library("ncdf4")
library("sf")
```

Mean annual temperature and precipitation are going to be useful for predicting soil carbon.
But this data isn't included in what we got from Tord. 
[The Climatic Research Unit](https://www.uea.ac.uk/groups-and-centres/climatic-research-unit)
has high-res global climate data available online [here](https://crudata.uea.ac.uk/cru/data/hrg/).

I've just downloaded the most recent data for monthly average daily mean temperature (degrees Celsius).
They also have precipitation (millimetres per month), but I'll start with temp first.

The data is stored as a NetCDF file.
I'm using this [this blog post](https://pjbartlein.github.io/REarthSysSci/netCDF.html) to figure out how to deal with that.

```{r}
temp_nc <- nc_open(here::here("data", "raw", "cru_ts4.07.2021.2022.tmp.dat.nc"))

print(temp_nc)

# get the variables out
lon <- ncvar_get(temp_nc,"lon")
lat <- ncvar_get(temp_nc,"lat")
tmp_array <- ncvar_get(temp_nc, "tmp")

# replace missing values with NA
fill_value <- ncatt_get(temp_nc, "tmp", attname = "_FillValue")

tmp_array <- tmp_array %>% 
  na_if(fill_value$value) 
  
```

```{r}
ncatt_get(temp_nc, "time", "units")

date <- ncvar_get(temp_nc,"time") %>% 
  as.Date(origin = "1900-1-1")

date
```

Ok, looks like we've got monthly average temperature for 2021 and 2022.

```{r}
dim(tmp_array)
```

The data is in a weird 3D raster format, with the third dimension being the date.
We should be able to visualise the data at a single time point by selecting one "slice".

```{r}
image(lon, lat, tmp_array[ , , 1], col = viridis_pal(option = "B")(10))
```

We need to reshape the data into a 2D data frame.

```{r}

make_2d <- function(data_slice, lon, lat) {
  lonlat <- as.matrix(expand.grid(lon, lat))
  data_vec <- as.vector(data_slice)
  cbind(lonlat, data_vec) %>% 
    data.frame() 
}

data_list <- apply(X = tmp_array, MARGIN = 3, FUN = make_2d, 
      lon = lon, lat = lat)

names(data_list) <- date

flat_data <- bind_rows(data_list, .id = "date") %>% 
  rename("tmp" = "data_vec",
         "lon" = "Var1",
         "lat" = "Var2") %>% 
    drop_na()

glimpse(flat_data)
```

Get MAT (mean annual temperature).

```{r}
mat_data <- flat_data %>% 
  mutate(year = substr(date, 1, 4)) %>% 
  group_by(lon, lat, year) %>% 
  summarise(mat = mean(tmp), .groups = "drop")

glimpse(mat_data)
```

Now we somehow need to match MAT to the NFI plots.

```{r warning=FALSE}
nfi_coords <- readxl::read_excel(
  here::here("data", "raw", "NFI plot coords NFI 2016-2020.xlsx"), 
  col_types = c("numeric", "text", "text", "numeric", "numeric", "numeric", 
                "numeric", "numeric", "numeric", "numeric", "numeric", 
                "numeric", "numeric")) %>% 
  select(-"...1") %>% 
  clean_names()

nfi_data <- readRDS(
  here::here("data", "derived", "ForManSims_RCP0_same_time.rds")
  ) %>% 
  filter(control_category_name == "Initial state") %>%
  mutate(description = str_replace_all(description, " ", "")) %>% 
  inner_join(nfi_coords)

```

Let's try `st_join()` with `st_nearest_feature`, 
It should match each plot to the nearest point in the climate data.

```{r}
# convert coordinates to sf objects

nfi_data_sf <- st_as_sf(x = nfi_data,                         
           coords = c("ost_wgs84", "nord_wgs84"),
           crs = "WGS84")

# will just use 2022 for now
mat_data_sf <- mat_data %>% 
  filter(year == "2022") %>% 
  st_as_sf(coords = c("lon", "lat"),
           crs = "WGS84")

nfi_mat <- st_join(
  nfi_data_sf,
  mat_data_sf,
  join = st_nearest_feature)
  
```

```{r, nfi-plots-2022-mat}
ggplot() + 
  geom_sf(data = nfi_mat, 
          aes(colour = mat), 
          alpha = 0.7) +
  scale_colour_viridis_c(option = "B") +
  ggtitle("Mean annual temperature, 2022")

```

Nice!

TO DO: precipitation and maybe get annual means for all years of period zero NFI plots.
