---
title: "Explore Heureka data"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

We have the first chunk of data from Tord, which is one fifth of the total data available.
Lets have a look at it!

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("janitor")
```

```{r}
tord_dat <- read.csv(
  here::here("data", "raw", "ForManSims_2016_RCP0_2.csv"), 
  sep = ";", skipNul = TRUE) %>% 
  clean_names()

glimpse(tord_dat)
```

How big is our actual sample size? 
First let's filter to only include plots that have been managed for both 'set aside' and 'BAU - no thinning'.

```{r}
tord_dat %>% 
  filter(control_category_name == "SetAside (Unmanaged)" |
           control_category_name == "BAU - NoThinning") ->  tord_dat_bau_sa

tord_dat_bau_sa %>% 
  group_by(description) %>% 
  summarise(n = n_distinct(control_category_name),
            .groups = "drop") %>% 
  filter(n == 2) %>% 
  select(description) -> plot_list

tord_dat_bau_sa %>% 
  filter(description %in% plot_list$description) -> plots_bau_sa
```

```{r}
plots_bau_sa %>% 
  filter(region == 5 & period == 20) %>% 
  ggplot(aes(x = age, y = total_soil_carbon, colour = control_category_name)) +
  geom_jitter(alpha = 0.5)
```

Becks says: "We can see that forest age is totally confounded with treatment. Of course! 
Clearcutting + replanting (BAU) resets the age of the forest. 
Unfortunately, this presents a challenge to implementing the potential outcome framework.
This is because the forests were different ages to start with (NFI input data), 
and so they will have been clearcut at different times, rendering different ages.
Because age is such a dominant predictor of soil carbon, it would need including in our RF models. 
But age is confounded with treatment.
To get around this problem - we can choose stands that were a similar starting age. 
So  that way, forest ‘age’ or ‘time since treatment’ are: totally confounded with treatment, and DO NOT VARY (so do not need specifying in the RF model).
Of course, this will sacrifice some sample size." 

Let's take a look at the age of the forests in the NFI input data.

```{r}
tord_dat %>% 
  filter(period == 0) %>% 
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 10)
```

If we filter to a starting age of 40, how many plots do we get?

```{r}
# get a list of plots with a starting age of 40
tord_dat %>% 
  filter(period == 0 & age == 40) %>% 
  select(description) %>% 
  distinct() -> plots_aged_40

# filter bau and set aside plots to only include plots with starting age of 40
plots_bau_sa %>% 
  filter(description %in% plots_aged_40$description) -> plots_bau_sa_40


n_distinct(plots_bau_sa_40$description)

```

Lets see what soil carbon is doing for a subset of plots.

```{r}
plots_bau_sa_40 %>% 
  filter(description == "2016 2019 4150" |
           description == "2016 1054 3180"|
           description == "2016 3002 2100"|
           description == "2016 5826 1031") %>% 
  ggplot(aes(x = period, y = total_soil_carbon, colour = control_category_name)) +
  geom_point(alpha = 0.7) +
  facet_wrap(~description, scales = "free_y")
```

Why are there so many replicates for BAU management?
Something to do with the `alternative_no` variable?

```{r}
plots_bau_sa_40 %>% 
  filter(description == "2016 2019 4150" |
           description == "2016 1054 3180"|
           description == "2016 3002 2100"|
           description == "2016 5826 1031") %>% 
  ggplot(aes(x = period, y = total_soil_carbon, 
             colour = as.factor(alternative_no))) +
  geom_point(alpha = 0.7) +
  geom_path() +
  facet_wrap(~description, scales = "free_y")
```

Tord said _"For each of the 32,000 plots, 
different management alternatives are simulated into the future (20(?) periods à 5 yrs). 
For each plot, these are numbered."_

It looks like the different management alternatives represented by `alternative_no` represent BAU starting at different times.
Ideally we want to pick an `alternative_no` where management happens at the same time period. 
Is there any structure to how `alternative_no` is numbered?

Try selecting the `alternative_no` which hits the highest peak `total_soil_carbon` for each plot.

```{r}
plots_bau_sa_40 %>% 
  filter(description == "2016 2019 4150" |
           description == "2016 1054 3180"|
           description == "2016 3002 2100"|
           description == "2016 5826 1031") %>% 
  group_by(description, alternative_no) %>% 
  summarise(max = max(total_soil_carbon)) %>% 
  slice_max(max, n = 1) %>% 
  inner_join(plots_bau_sa_40) %>% 
  ggplot(aes(x = period, y = total_soil_carbon, 
             colour = as.factor(alternative_no))) +
  geom_point(alpha = 0.7) +
  geom_path() +
  facet_wrap(~description, scales = "free_y")
```
Now try selecting `alternative_no`s which hit their peak in the same `period`.

```{r}
plots_bau_sa_40 %>% 
  filter(description == "2016 2019 4150" |
           description == "2016 1054 3180"|
           description == "2016 3002 2100"|
           description == "2016 5826 1031") %>% 
  filter(period == 10) %>% 
  group_by(description, alternative_no, .drop = FALSE) %>% 
  summarise(max = max(total_soil_carbon)) %>% 
  slice_max(max, n = 1) %>% 
  inner_join(plots_bau_sa_40) %>% 
  ggplot(aes(x = period, y = total_soil_carbon, 
             colour = as.factor(alternative_no))) +
  geom_point(alpha = 0.7) +
  geom_path() +
  facet_wrap(~description, scales = "free_y")
```

I guess this is the most useful for putting into the model?

## Peat

```{r}
plots_bau_sa_40 %>% 
  filter(peat == 1) %>% 
  distinct(description)
```

There are just four plots in this subset where `peat == 1`.

```{r}
plots_bau_sa_40 %>% 
  filter(peat == 1) %>% 
  ggplot(aes(x = period, y = total_soil_carbon, 
             colour = as.factor(alternative_no))) +
  geom_point(alpha = 0.7) +
  geom_path() +
  facet_wrap(~description, scales = "free_y")
```

[Heureka wiki](https://www.heurekaslu.se/wiki/Carbon_sequestration): __On ditched organic soils (peat), soil carbon decreases over time (using emission factors), while the soil carbon stock in unditched peatland remains constant (default 896 ton C/ha) over time.__

This makes sense with those figures. 
We probably don't want to use plots with peat soils.
