---
title: "Trial new geographical test area argument in function"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

We found a geographical area containing plots with lots of covariate overlap to use as testing data. Previously we were using random 200 plots but ideally people wouldn’t predict outside of parameter space. 

I've implemented a new argument in the `fit_metalearner` function 
which allows you to specify if you want to use a random sample of plots for testing
or use the geographically defined "square" of plots.

I've also modified the `assign_treatment` function to __not__ assign a treatment
to the plots in the square. 
I now randomly assign treatments for these plots in the `fit_metalearner` 
function so that we get an even split of 0 and 1.

Will be testing these out below.

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("maps")
library("patchwork")
```

```{r}
set.seed(123)

# get my functions
function_dir <- list.files(here::here("code", "functions"),
                           full.names = TRUE)

sapply(function_dir, source)
```

```{r}
data_clean <-
  readRDS(here::here(
    "data", "derived", "ForManSims_RCP0_same_time_clim_squ.rds"
    )) 
```

```{r}
data_assigned <- assign_treatment(df = data_clean,
                 assignment = "random",
                 proportion_not_treated = 0.5)
```

```{r}
model_out_squ <- fit_metalearner(
     df = data_assigned,
     learner = "s",
     n_train = 1400,
     var_omit = FALSE,
     random_test_plots = FALSE)

model_out_rand <- fit_metalearner(
     df = data_assigned,
     learner = "s",
     n_train = 1400,
     var_omit = FALSE,
     random_test_plots = TRUE)
```

```{r}
plot_real_pred_s <- function(out) {
  out %>% 
    ggplot() +
    geom_hline(yintercept = 0, colour = "grey", linetype = 2) +
    geom_vline(xintercept = 0, colour = "grey", linetype = 2) +
    geom_point(aes(x = cate_real, y = cate_s_learn, colour = abs(diff))) +
    geom_abline(intercept = 0, slope = 1, colour = "blue") +
    scale_color_gradient(low = "lightblue", high = "red3",
                         limits = c(0, 30)) +
    xlim(-30, 30) +
    ylim(-30, 30)
}

plot_real_pred_s(model_out_rand) +
  plot_real_pred_s(model_out_squ) +
  plot_layout(guides = "collect")

```

```{r}
model_out_rand %>% 
  summarise(median_abs_error = median(abs(diff))) 
```

```{r}
model_out_squ %>% 
  summarise(median_abs_error = median(abs(diff))) 
```
