---
title: "Tweaks for treatment assignment ~ region correlation"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

Treatment assignment correlated with region - needs to be bigger percentages! 
Currently we are using: 12% 8.5%, 7%, 7%, 12.5%, 12.5% - 
these percentages were based off of real data but 
they might be too similar to see much difference between them and 
beacuse they are so small we aren't getting any non-treated plots when the sample size is small.

I've updated the function `assign_treatment()` to use 20%, 40%, 60%, 80% - 
ordered or random with respect to latitude.

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("maps")
library("patchwork")
```

```{r}
set.seed(123)

# get my functions
function_dir <- list.files(here::here("code", "functions"),
                           full.names = TRUE)

sapply(function_dir, source)
```

```{r}
clean_data <-
  readRDS(here::here("data", "derived", "ForManSims_RCP0_same_time_clim.rds")) 
```

```{r}
keys <- expand.grid(
  assignment = c("random", "blocked_ordered", "blocked_random", "correlated_altitude")
  )

assigned_data <- purrr::map(.f = assign_treatment, 
                     .x = keys$assignment,  
                     df = clean_data)

keys %>% 
  mutate(df = assigned_data) -> keys_assigned

```

```{r}
make_map <- function(df, name) {
  df %>% 
  filter(period == 0) %>%
  ggplot(aes(ost_wgs84, nord_wgs84, colour = as.factor(tr))) +
  borders("world", regions = "sweden") +
  geom_point(alpha = 0.5, shape = 16) +
  coord_quickmap() +
  theme_void() +
  theme(legend.title=element_blank()) +
  ggtitle(name)
}

plot_list <- map2(.x = keys_assigned$df,
     .y = keys_assigned$assignment,
     .f = make_map) 

wrap_plots(plot_list, nrow = 1)
```

```{r warning=FALSE}
model_out_s <- purrr::map(.x = keys_assigned$df,
     .f = fit_metalearner,
     learner = "s",
     n_train = 1600,
     var_omit = FALSE)

keys_assigned %>% 
  mutate(out = model_out_s) -> keys_out_s

plot_real_pred_s <- function(out, assignment) {
  out %>% 
    ggplot() +
    geom_hline(yintercept = 0, colour = "grey", linetype = 2) +
    geom_vline(xintercept = 0, colour = "grey", linetype = 2) +
    geom_point(aes(x = cate_real, y = cate_s_learn, colour = abs(diff))) +
    geom_abline(intercept = 0, slope = 1, colour = "blue") +
    scale_color_gradient(low = "lightblue", high = "red3",
                         limits = c(0, 35)) +
    xlim(-35, 35) +
    ylim(-35, 35) +
    ggtitle(assignment)
}

plot_list_ite_s <- map2(.x = keys_out_s$out,
     .y = keys_out_s$assignment,
     .f = plot_real_pred_s) 

wrap_plots(plot_list_ite_s, nrow = 2) + 
  plot_layout(guides = "collect") +
  plot_annotation(title = "S-learner")
```

```{r warning=FALSE, echo=FALSE}
model_out_t <- purrr::map(.x = keys_assigned$df,
     .f = fit_metalearner,
     learner = "t",
     n_train = 1600,
     var_omit = FALSE)

keys_assigned %>% 
  mutate(out = model_out_t) -> keys_out_t

plot_real_pred_t <- function(out, assignment) {
  out %>% 
    ggplot() +
    geom_hline(yintercept = 0, colour = "grey", linetype = 2) +
    geom_vline(xintercept = 0, colour = "grey", linetype = 2) +
    geom_point(aes(x = cate_real, y = cate_t_learn, colour = abs(diff))) +
    geom_abline(intercept = 0, slope = 1, colour = "blue") +
    scale_color_gradient(low = "lightblue", high = "red3",
                         limits = c(0, 35)) +
    xlim(-35, 35) +
    ylim(-35, 35) +
    ggtitle(assignment)
}

plot_list_ite_t <- map2(.x = keys_out_t$out,
     .y = keys_out_t$assignment,
     .f = plot_real_pred_t) 

wrap_plots(plot_list_ite_t, nrow = 2) + 
  plot_layout(guides = "collect") +
  plot_annotation(title = "T-learner")
```

```{r warning=FALSE, echo=FALSE}
model_out_x <- purrr::map(.x = keys_assigned$df,
     .f = fit_metalearner,
     learner = "x",
     n_train = 1600,
     var_omit = FALSE)

keys_assigned %>% 
  mutate(out = model_out_x) -> keys_out_x

plot_real_pred_x <- function(out, assignment) {
  out %>% 
    ggplot() +
    geom_hline(yintercept = 0, colour = "grey", linetype = 2) +
    geom_vline(xintercept = 0, colour = "grey", linetype = 2) +
    geom_point(aes(x = cate_real, y = cate_x_learn, colour = abs(diff))) +
    geom_abline(intercept = 0, slope = 1, colour = "blue") +
    scale_color_gradient(low = "lightblue", high = "red3",
                         limits = c(0, 35)) +
    xlim(-35, 35) +
    ylim(-35, 35) +
    ggtitle(assignment)
}

plot_list_ite_x <- map2(.x = keys_out_x$out,
     .y = keys_out_x$assignment,
     .f = plot_real_pred_x) 

wrap_plots(plot_list_ite_x, nrow = 2) + 
  plot_layout(guides = "collect") +
  plot_annotation(title = "X-learner")
```
