---
title: "Comparing measures of predictive accuracy"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("patchwork")
```

```{r}
results <- readRDS(here::here("data", "derived", "results.rds")) %>% 
  filter(n_train != 31) 
```

```{r fig.height=10, fig.width=20, warning=FALSE}
results %>% 
  ggplot(aes(x = n_train, y = rmse, colour = learner)) +
  stat_summary(fun = median, geom = "line", linewidth = 1) +
  stat_summary(fun = median, size = 1, fill = "white", shape = 21, stroke = 1.5) +
  scale_colour_manual(values = c("#009E73", "#E69F00", "#0072B2")) +
  scale_x_continuous(breaks = c(0, 62, 125, 250, 500, 1000)) +
  xlab("Sample size") +
  ylab("RMSE") +
  theme_classic(base_size = 15) +
  
  ggplot(results, aes(x = n_train, y = rsq, colour = learner)) +
  stat_summary(fun = median, geom = "line", linewidth = 1) +
  stat_summary(fun = median, size = 1, fill = "white", shape = 21, stroke = 1.5) +
  scale_colour_manual(values = c("#009E73", "#E69F00", "#0072B2")) +
  scale_x_continuous(breaks = c(0, 62, 125, 250, 500, 1000)) +
  xlab("Sample size") +
  ylab("R squared") +
  theme_classic(base_size = 15) +
  
  ggplot(results, aes(x = n_train, y = median_error, colour = learner)) +
  stat_summary(fun = median, geom = "line", linewidth = 1) +
  stat_summary(fun = median, size = 1, fill = "white", shape = 21, stroke = 1.5) +
  scale_colour_manual(values = c("#009E73", "#E69F00", "#0072B2")) +
  scale_x_continuous(breaks = c(0, 62, 125, 250, 500, 1000)) +
  xlab("Sample size") +
  ylab("Median error") +
  theme_classic(base_size = 15) +
  
  ggplot(results, aes(x = n_train, y = mean_error, colour = learner)) +
  stat_summary(fun = median, geom = "line", linewidth = 1) +
  stat_summary(fun = median, size = 1, fill = "white", shape = 21, stroke = 1.5) +
  scale_colour_manual(values = c("#009E73", "#E69F00", "#0072B2")) +
  scale_x_continuous(breaks = c(0, 62, 125, 250, 500, 1000)) +
  xlab("Sample size") +
  ylab("Mean error") +
  theme_classic(base_size = 15) +
  
  plot_layout(guides = "collect")
  
```

RMSE: A metric that tells us how far apart the predicted values are from the observed values in a dataset, on average. The lower the RMSE, the better a model fits a dataset.

R2: A metric that tells us the proportion of the variance in the response variable of a regression model that can be explained by the predictor variables. This value ranges from 0 to 1. The higher the R2 value, the better a model fits a dataset.
