---
title: "Find a region of covariate overlap"
author: '`r Sys.getenv("USER")`'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

Find a geographical area of covariate overlap to use as testing data -
we want the area to be geographically distinct but environmentally nested /
nested in covariate space.
Currently using random 200 plots but ideally in the real world people wouldn’t 
predict outside of parameter space. 

```{r packages, message=FALSE, warning=FALSE}
library("tidyverse")
library("here")
library("FactoMineR")
library("factoextra")
library("ggmap")
library("maps")
library("sf")
library("patchwork")
library("GGally")

clean_data <-
  readRDS(here::here("data", "derived", "ForManSims_RCP0_same_time_clim.rds")) %>% 
  filter(period == 0)
```

## Normalise data

```{r}
feat_list <- c("total_soil_carbon", "altitude",
                   "mat_5yr", "map_5yr", "ditch", "no_of_stems", "volume_pine",
                   "volume_spruce", "volume_birch", "volume_aspen",
                   "volume_oak", "volume_beech", "soil_moist_code",
                   "volume_southern_broadleaf", "volume_larch")

corr_matrix <- clean_data %>% 
  column_to_rownames(var = "description") %>%  
  select(all_of(feat_list)) %>% 
  scale() 
```

## Make PCA

```{r}
# compute pca
data_pca <- princomp(corr_matrix)

# see results 
summary(data_pca)
```

## Plot

```{r}
fviz_pca_var(data_pca)
```

### Latitude

```{r fig.width = 10, fig.height = 5}
fviz_mca_ind(data_pca, 
             geom = "point", 
             col.ind = clean_data$nord_wgs84) + 
  scale_colour_viridis_c() +
  
  ggplot(clean_data, aes(ost_wgs84, nord_wgs84, colour = nord_wgs84)) +
  borders("world", regions = "sweden") +
  geom_point() +
  scale_color_viridis_c() +
  coord_quickmap() +
  theme(legend.position = "none")
```

### Longitude

```{r fig.width = 10, fig.height = 5}
fviz_mca_ind(data_pca, 
             geom = "point", 
             col.ind = clean_data$ost_wgs84)+ 
  scale_colour_viridis_c() +
  
  ggplot(clean_data, aes(ost_wgs84, nord_wgs84, colour = ost_wgs84)) +
  borders("world", regions = "sweden") +
  geom_point() +
  scale_color_viridis_c() +
  coord_quickmap() +
  theme(legend.position = "none")
```

Looks like somewhere in the middle of Sweden would be the most homogeneous 
with the rest of the country.

### Region

```{r fig.width = 10, fig.height = 5}
fviz_mca_ind(data_pca, 
             geom = "point", alpha = 0.7,
             col.ind = as.factor(clean_data$region)) +
  
  ggplot(clean_data, aes(ost_wgs84, nord_wgs84, colour = as.factor(region))) +
  borders("world", regions = "sweden") +
  geom_point(alpha = 0.7) +
  coord_quickmap() +
  theme(legend.position = "none")
```

## Create grid

```{r}
# make data a sf object
data_sf <- st_as_sf(clean_data, 
                    coords = c("ost_wgs84", "nord_wgs84"),
                    crs = "WGS84")

# project the points
data_sf %>% 
  st_transform(crs = st_crs(3152)) -> data_projected

# make grid
fishnet <- st_make_grid(
  data_projected,
  cellsize = c(140000, 130000), # units are meters
  what = "polygons",
  square = TRUE,
  crs = st_crs(3152))

# plot
ggplot(data_projected) +
  geom_sf(aes(col = as.factor(region)), alpha = 0.6, shape = 16) +
  geom_sf(data = fishnet, fill  = NA)
```

How many plots in each rectangle? We can have up to 200.

```{r warning=FALSE}
# convert polygons to sf object and add id column
fishnet %>%
	st_sf(crs = st_crs(3152)) %>%
	mutate(net_id = row_number()) -> fishnet_sf

# calculate which plots are in which squares
joined <- st_intersection(data_projected, fishnet_sf)

joined %>% 
  group_by(net_id) %>% 
  summarise(n()) %>% 
  slice_max(order_by = `n()`, n = 10)
```

Let's compare the top 4.

```{r}
joined %>% 
  filter(net_id == 22 | net_id == 17 | net_id == 7 | net_id == 27) %>%  
  ggplot() +
  geom_sf(aes(col = as.factor(region)), alpha = 0.6, shape = 16) +
  facet_wrap(~net_id, nrow = 1)
```

```{r message=FALSE}
clean_data %>% 
  left_join(joined) -> data_nets

data_nets %>% 
  mutate(net_col = case_when(net_id == 22 ~ "net_22",
                            net_id == 17 ~ "net_17",
                             net_id == 7 ~ "net_7",
                             net_id == 27 ~ "net_27",
                              .default = "other")) -> nets_test

fviz_mca_ind(data_pca, 
             geom = "point", alpha = 0.6, shape = 16,
             col.ind = as.factor(nets_test$net_col), 
             palette = c("red", "blue", "forestgreen", "orange", "lightgrey")) 
```

```{r echo=FALSE}
data_nets %>% 
  mutate(net_col = case_when(net_id == 17 ~ "net_17",
                              .default = "other")) -> nets_test_17

fviz_mca_ind(data_pca, 
             geom = "point", alpha = 0.5, shape = 16,
             col.ind = as.factor(nets_test_17$net_col), addEllipses = TRUE,
             palette = c("red", "lightgrey")) -> p1

data_nets %>% 
  mutate(net_col = case_when(net_id == 22 ~ "net_22",
                              .default = "other")) -> nets_test_22

fviz_mca_ind(data_pca, 
             geom = "point", alpha = 0.5, shape = 16,
             col.ind = as.factor(nets_test_22$net_col), addEllipses = TRUE,
             palette = c("blue", "lightgrey")) -> p2

data_nets %>% 
  mutate(net_col = case_when(net_id == 27 ~ "net_27",
                              .default = "other")) -> nets_test_27

fviz_mca_ind(data_pca, 
             geom = "point", alpha = 0.5, shape = 16,
             col.ind = as.factor(nets_test_27$net_col), addEllipses = TRUE,
             palette = c("forestgreen", "lightgrey")) -> p3

data_nets %>% 
  mutate(net_col = case_when(net_id == 7 ~ "net_7",
                              .default = "other")) -> nets_test_7

fviz_mca_ind(data_pca, 
             geom = "point", alpha = 0.5, shape = 16,
             col.ind = as.factor(nets_test_7$net_col), addEllipses = TRUE,
             palette = c("orange", "lightgrey")) -> p4

(p1 + p2) /
  (p3 + p4)
```

It looks like `net_22` is the most nested within the covariate space. 
When we remove these points 
what effect do they have on the distribution of our covariates?

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
data_nets %>% 
  select(total_soil_carbon, altitude,
         mat_5yr, map_5yr, ditch, no_of_stems, volume_pine,
         volume_spruce, volume_birch, volume_aspen,
         volume_oak, volume_beech, soil_moist_code,
         volume_southern_broadleaf, volume_larch) %>% 
  mutate(ditch = as.ordered(ditch), 
         ditch = as.ordered(soil_moist_code)) %>% 
  ggpairs(progress = FALSE,
          mapping = aes(alpha = 0.5)) +
  theme_classic(base_size = 7) 
```

```{r fig.height=10, fig.width=10, message=FALSE, warning=FALSE}
data_nets %>% 
  mutate(net_col = case_when(net_id == 22 ~ "net_22",
                              .default = "other")) %>% 
  select(net_col, total_soil_carbon, altitude,
         mat_5yr, map_5yr, ditch, no_of_stems, volume_pine,
         volume_spruce, volume_birch, volume_aspen,
         volume_oak, volume_beech, soil_moist_code,
         volume_southern_broadleaf, volume_larch) %>% 
  mutate(ditch = as.ordered(ditch), 
         ditch = as.ordered(soil_moist_code)) %>% 
  ggpairs(progress = FALSE,
          mapping = aes(colour = as.factor(net_col), alpha = 0.5)) +
  theme_classic(base_size = 7)  
```

In the above, pink is `net_22`. 
It looks good - pink is always within the range of blue,
although shifted a bit for altitude..

```{r}
# make data a sf object
data_sf_net <- data_nets %>% 
  mutate(net_col = case_when(net_id == 22 ~ "net_22",
                              .default = "other")) %>% 
  st_as_sf(coords = c("ost_wgs84", "nord_wgs84"),
           crs = "WGS84")

# project the points
data_sf_net %>% 
  st_transform(crs = st_crs(3152)) -> data_projected_net

# plot
ggplot(data_projected_net) +
  geom_sf(aes(col = as.factor(net_col)), alpha = 0.6, shape = 16) 
```
