---
title: "Check for autocorrelation"
author: 'eleanorjackson'
date: '`r format(Sys.time(), "%d %B, %Y")`' 
always_allow_html: true
output: 
  github_document:
    keep_html: true
---
	
```{r setup, include = FALSE}
file_name <- rstudioapi::getSourceEditorContext()$path

knitr::opts_chunk$set(
  fig.path =
    paste0("figures/", sub("\\.Rmd$", "", basename(file_name)), "/", sep = "")
)

ggplot2::theme_set(ggplot2::theme_classic(base_size = 15))
```

```{r}
library("tidyverse")
library("here")
library("patchwork")
library("spdep")
```

Check for spatial autocorrelation in RMSE:

We have errors plotted in Fig 3 - R1 has said they look spatially autocorrelated. 
It’s worth checking spatial autocorrelation explicitly? Perhaps you could do this for conditions: n = 1000, treatment ass = random, imbalance = 0.5, etc

coords <- as.matrix(d[, c("x", "y")]) nb <- knn2nb(knearneigh(coords, k = 8)) # k nearest neighbours lw <- nb2listw(nb, style = "W") moran.test(d$error, lw) # I, expected I, p-value 

Compute Moran's I per run and compare its distribution across our three test-location conditions?


```{r}
all_runs <- 
  readRDS(here::here("data", "derived", "all_runs.rds")) %>% 
  filter(n_train == 1000, 
         assignment == "random", 
         test_plot_location == "stratified",
         var_omit == "none",
         prop_not_treated == 0.5) 

one_run <- 
  all_runs$df_out[[1]]
```

```{r warning=FALSE}
nfi_coords <- readxl::read_excel(
  here::here("data", "raw", "NFI_plot_coords_NFI_2016-2020.xlsx"),
  col_types = c("numeric", "text", "text","numeric","numeric","numeric",
                "numeric","numeric","numeric","numeric","numeric","numeric","numeric"),
  progress = FALSE
) %>%
  select(Description, Ost_WGS84, Nord_WGS84) %>%
  janitor::clean_names()
```


```{r}
clean_data <-
  readRDS(here::here("data", "derived", "ForManSims_RCP0_same_time_clim_squ.rds"))
```

```{r}
one_run %>%
    mutate(Error = cate_pred - cate_real) %>%
    left_join(nfi_coords, by = join_by(description)) -> plot_dat
```

```{r}
coords <- as.matrix(plot_dat[, c("ost_wgs84", "nord_wgs84")]) 

nb <- knn2nb(knearneigh(coords, k = round(sqrt(nrow(plot_dat))))) # k nearest

lw <- nb2listw(nb, style = "W") 

moran.test(plot_dat$Error, lw) # I, expected I, p-value 

moran.test(plot_dat$Error, lw)$estimate[[1]] # return this value
```

```{r}
all_runs <- all_runs %>%
  mutate(moran_i = purrr::map(
    .x = df_out,
    .f = ~ moran_i_vec(xcoord = .x$ost_wgs84,
                       ycoord = .x$nord_wgs84,
                       value = .x$error)
  )) %>%
  unnest(moran_i)

```

